Sketch Recogniton using Particle Swarm Algorithm

### 2009-02-18 draft/ChapterIntroduction.tex
before removing the comment

Sketchs here stand for any hand drawn drawing the user draws using digital stylus or similar device. Normally, sketch are composed of smaller segments we call strokes. The time between pen-up and pen-down events identify each stroke.  The strokes are mainly the path of points, which the user path through between the pen-down and pen-up events while using the stylus or similar device.

@@

 Paper and Pencil are preferred in design as they offer a remarkable ease in editing and altering the primary rough sketches. Current CAD systems do not give the designer the ease of altering the early designs. Designer had to sketch primary designs on paper and pencil next, the finally design is transferred into computer using a menu CAD system. The process is time consuming and tiring. The need for fast design and development process is driven by the fast growing industries and technology.

 A sketch recognition and understanding system will provide a powerful tool to the designers to draw sketch as they usually draw using paper and pencil with the advantages of computer program. Designer needs a normal sketching environment that will provide him facilities in paper interface and the modifying and simulation that is in current CAD systems.

@@

\section{Research Scope}
%rewirte this part.............................
%  research will invistgated till sketch recognition in domain of uml diagrams. 
%This research will include investigating the best method to recognize sketches.  The scope of the research will include recognizing various sketches. The methods used in this research are comparing current systems and enhancing them and adobpiting them into simple geometrical symbols.  Using well established AI methods specially \textit{Swarm Intelligence} to solve the problem.

\section{Problem Definition}
\label{sec:ProblemsAndChallenges}

%Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition. The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information. In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments. In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.

@@

The research in the area had faced few problems that will need to be handled to implement an entirely sketch understanding system. The user draws strokes in a spontaneous and ambiguous manner that intensifies the challenges. The first problem is segmenting strokes from the way the user draw it into primitive geometric. Ink parsing is the problem dealing with clustering interpolated strokes and grouping them into symbols. After that, recognizing the symbols from the space the system can identify is the final problem. All those problems must be solved on a real time constrain to maintain an online sketching system if it was a requirement.

The next sections will explain and investigate each of those problems in details.

\subsection{Segmentations}
\label{sec:Segmentations}

Segmentation is dividing the strokes draws into the geometric primitive's as lines, arcs, and ellipses. The strokes as mentioned before are the path of points between pen-down and pen-up events while using a stylus. The processing of those points is done after that to firstly to detect the critical points or vertices. After that the segment is classified as either line, arc, ellipse or circle. The main problem in this process is in detecting the critical and vertices which are points divide strokes into lines and arc. The user draw the lines and arc in an undefined manner which make it difficult for the program to deicide where the user had finish a line and started a new arc or vice verse; Fig [\ref {fig:strokeinterpolation}].

@@

\subsection{Ink parsing}
\label{sec:InkParsing}

 The clustering of segmented strokes into group of strokes before passing into the next stage this process is called ink parsing. It is simply deciding when a symbols ends and a new symbol begins.  Ink parsing is most important and difficult challenge as grouping of strokes can be done in many ways and configuration, which increase exponentially with the number of strokes. In addition, the ink parsing will affect the next steps of recognizing. The user does not draw the symbol in same order. Moreover, users can draw a sequence of strokes in symbols then leave it to draw another then return again to add segment to symbols so temporal data is helpful but not conclusive.  The user may also draw the same symbols using different orders and methods for example the user can draw a rectangle using single stroke or using two, three or even four strokes. All those factors gather with the sloppy drawing of the user to increase the complexity of the problem.

\subsection{Ambiguities}
\label{sec:Ambiguities}

The major problem of sketch understanding is the high level of ambiguities in the drawing. The user draws strokes that the computer cannot decide if it is arc or line. The nature of hand drawn sketch it self is highly ambiguous even for human eye you may see two drawn you cannot decide if it is square or rectangle.
 The ambiguities in sketch understanding are found at many levels in the system. Firstly, at the segmentation phase where the same user stroke can be interpreted as a ellipse and a circle. Then the symbol recognition as a symbol can have more than one interpolation. Due to the user sloppy drawing the symbols can have more than one interpolation which varies with the domain drawn in it similar problem was faced in handwriting recognition see Fig [\ref{fig:VorU}] but it was handled with knowing information about the context information which may not be the case in sketch recognition.

@@

The ambiguity problem must be first recognized ambiguities or varicose interpolations of same symbol. After that, resolve those ambiguities is detecting the best correct interpolation of the symbol.  To solve this problem the program must be able to detect the amount of information needed to decide which interpolation of symbols is the best one.

>>>>

Sketchs here stand for any hand drawn drawing the user draws using digital stylus or similar device. Normally, sketchs are composed of a set of strokes. The time between pen-up and pen-down events identify each stroke. The strokes are mainly the path of points, which the user path through between the pen-down and pen-up events while using the stylus or similar device.

@@

 Paper and Pencil are preferred in design as they offer a remarkable ease in editing and altering the primary rough sketches. Current CAD systems do not give the designer the ease of altering the early designs. Designer had to sketch primary designs on paper and pencil next, the final design is transferred into computer using a CAD system menus. The process is time consuming and tiring. The need for fast design and development process is driven by the fast growing industries and technology.

 A sketch recognition and understanding system will provide a powerful tool to the designers to draw sketch as they usually draw using paper and pencil with the advantages of a computer program. Designer needs a normal sketching environment that will provide him facilities in paper interface and the modifying and simulation that is in current CAD systems.

@@

\section{Problem Definition and challenges}
\label{sec:ProblemsAndChallenges}

@@

Sketch recognition is divided into three steps preprocessing, segmentation and symbol recognition. The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information. In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments. In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.

The research in the area had faced few problems that will need to be handled to implement an entirely sketch recognition system. The user draws strokes in a spontaneous and ambiguous manner that intensifies the challenges. One of the problems is segmenting strokes from the way the user draw it into primitive geometric. Another major problem is the high level of ambiguities in the drawing. The next sections discripe each problem in details 
\subsection {Segmentation}
Segmentation is dividing the strokes draws into the geometric primitive's as lines, arcs, and ellipses. The strokes as mentioned before are the path of points between pen-down and pen-up events while using a stylus. The processing of those points is done after that to firstly to detect the critical points or vertices. After that the segment is classified as either line, arc, ellipse or circle. The main problem in this process is in detecting the critical and vertices which are points divide strokes into lines and arc. The user draw the lines and arc in an undefined manner which make it difficult for the program to deicide where the user had finish a line and started a new arc or vice verse; Fig [\ref {fig:strokeinterpolation}].

@@

\subsection{Ambiguity}
The user draws strokes that the computer cannot decide if it is arc or line. The nature of hand drawn sketch it self is highly ambiguous even for human eye you may see two drawn you cannot decide if it is square or rectangle.

 The ambiguities in sketch understanding are found at many levels in the system. Firstly, at the segmentation phase where the same user stroke can be interpreted as a ellipse and a circle. Symbol ambiguity is antoher problem as a symbol can have more than one interpolation. Due to the user sloppy drawing the symbols can have more than one interpolation which varies with the domain drawn in it similar problem was faced in handwriting recognition see Fig [\ref{fig:VorU}] but it was handled with knowing information about the context information which may not be the case in sketch recognition.

@@

\section{Research Scope}
%rewirte this part.............................
%  research will invistgated till sketch recognition in domain of uml diagrams. 
%This research will include investigating the best method to recognize sketches.  The scope of the research will include recognizing various sketches. The methods used in this research are comparing current systems and enhancing them and adobpiting them into simple geometrical symbols.  Using well established AI methods specially \textit{Swarm Intelligence} to solve the problem. 
 This research investigats the best method to recognize sketches. The scope of the research includes recognizing various symbols drawn using single and multipule strokes. The research foucs on using various Particle Swarm Optimization (PSO) methods to segments strokes into goemetrical primitives. Sketch recognition is achieved by  Support Vector Machines (SVM) classifier. A various set of features are investigated to reach the best set of features.
